Determining Frequentist Confidence Limits Using a Directed Parameter Space Search

### 2014-08-04 aps4p5/drafts/draft_current.tex
new draft up to date; just need to wait for the 1280.7 run to finish

{\bf %marking new text
Readers should be aware that, while MCMC can be analytically shown to
asymptotically sample the Bayesian posterior, APS is designed to discover
Frequentist confidence limits purely through exploration, i.e. there is no {\it a
priori} motivation behind the APS algorithm.  Our only demonstration of APS'
efficacy is empirical.  It has done well on the problems we have tested it on,
though we have tried to choose those problems such that they are representative of
problems likely to be encountered by the community.  Readers who require
theoretically-motivated results will likely need to resort to a sampling routine
like MCMC to derive their final constraints.  We hope to show, however, in Section
\ref{sec:toy}, that even in these use cases, APS can provide a valuable service by
identifying the general regions of parameter space over which the Bayesian
posterior needs to be integrated, i.e. by guiding researchers in the process of setting
priors for a sampling algorithm, a problem which has posed significant problems
for MCMC in the case of multi-modal likelihood functions.
} %marking new text

@@

At this point, each $\vec{a}$ in $\{\hat{A}\}$ should have a corresponding symmetric hyperbox
surrounding it described by the corresponding 
$\vec{a}_\text{max}(\vec{a})$ and $\vec{a}_\text{min}(\vec{a})$.  
To integrate the posterior probability density, we will assume that all of the
parameter space points in each hyperbox correspond have $\chi^2=\chi^2(\vec{a})$,
{\bf %marking new text
i.e. the $\chi^2$ discovered by APS for the point at the center of the hyperbox.
} %marking new text
The posterior probability associated with each box is then
\begin{eqnarray}
P(\vec{a})&=&\frac{V(\vec{a})}{P_\text{tot}}\times\exp\left[-\chi^2(\vec{a})/2\right]
\label{eqn:posterior}\\
P_\text{tot}&\equiv&\sum_{\vec{a}} V(\vec{a})\times\exp\left[-\chi^2(\vec{a})/2\right]
\end{eqnarray}
where $V(\vec{a})$ is the hypervolume of the hyperbox surrounding $\vec{a}$.
To plot the $(1-\alpha)\%$ credible limit, one need only plot the points in hyperspace that
enclose the lowest-$\chi^2$ hyperboxes containing $(1-\alpha)\%$ of the total probability.
We test this algorithm on a function with an exactly known credible limit
in Section \ref{sec:toy} (see Figures \ref{fig:toyBayes2}-\ref{fig:toyBayes4}) 
and on a real, physical dataset in Section \ref{sec:wmap} (see Figure \ref{fig:contours_bayes}).  
In  Section \ref{sec:toy}, we find that our results agree with the known credible limit.
{\bf %marking new text
In Section \ref{sec:wmap} we find that APS correctly identifies the region and general
shape of the Bayesian credible limits, however, with significant noise.  The Bayesian
credible limits found by APS are systematically larger than the Bayesian credible
limits found by MCMC.  This motivates our claim in the introduction that APS may be
principally useful for identifying regions of high posterior density on largely unknown
likelihood functions, allowing researchers to more shrewdly set thep priors for sampling
algorithms like MCMC.
}%marking nex text

@@

In Figure \ref{fig:contours_bayes}, we plot the 95\% Bayesian credible limit discovered
by APS after 50,000 evalutations of $\chi^2$ and determined by the method outlined in
Section \ref{sec:bayes}.  Unfortunately, we do not see here the convergence to the true
credible limits that we say in Figures \ref{fig:toyBayes2}-\ref{fig:toyBayes4}.  While
it does appear that APS has correctly identified the locations and general shapes of the
true credible limits, it has, in most cases, drawn credible limits that are larger than
those found by MCMC.  This should not be surprising.  APS was designed with Frequentist
confidence limits in mind, and Figures \ref{fig:contours_gauss}-\ref{fig:contours_br}
seem much better converged than Figure \ref{fig:contours_bayes}.  However, it does seem
that APS might be useful in cases where little prior knowledge is known about $\vec{\theta}$
and researchers need to be confident that they have discovered all of the low-$\chi^2$
regions in a particular parameter space.  Once APS has been used to identify rough
Bayesian credible limits, more detailed algorithms (like MCMC) can be run using the prior
knowledge gained by APS to refine the credible limits.

@@

In Section \ref{sec:toy} we indicated that, consistent with the findings
of Bryan {\it et al}. (2007), APS is more robust
than MCMC against multi-modal likelihood functions (at least on the idealized
cartoon $\vec{\theta}\rightarrow\chi^2$ functions considered).
In Section \ref{sec:wmap} we showed that on realistic data, APS
is able to yield Frequentist confidence limits that are comparable to
Bayesian credible limits with significantly fewer calls to $\vec{\theta}\rightarrow\chi^2$
than MCMC (recall that APS required only 50,000 $\chi^2$ evaluations to yield Figures
\ref{fig:contours_gauss}-\ref{fig:contours_br}).  However, the Bayesian credible
limits yielded by APS remain rough and imprecise.  We believe that this performance
can still be of use to the community, especially when trying to evaluate proof-of-concept
parameter constraints (perhaps for forecasting future parameter constraints)
with very expensive $\vec{\theta}\rightarrow\chi^2$ functions on poorly-understood
parameter spaces.  Certainly, for researchers interested in plotting Frequentist confidence limits,
APS is a viable option.

>>>>

{\bf %marking new text
Readers should be aware that, while MCMC can be analytically shown to
asymptotically sample the Bayesian posterior, APS is designed to discover
Frequentist confidence limits purely through exploration, i.e. there is no {\it a
priori} motivation behind the APS algorithm.  Our only demonstration of APS'
efficacy is empirical.  It has done well on the problems we have tested it on,
though we have tried to choose those problems such that they are representative of
problems likely to be encountered by the community.  Readers who require
theoretically-motivated results will likely need to resort to a sampling routine
like MCMC to derive their final constraints.  We hope to show, however, in Section
\ref{sec:toy}, that even in these use cases, APS can provide a valuable service by
identifying the general regions of parameter space over which the Bayesian
posterior needs to be integrated, i.e. by guiding users in the process of setting
priors and proposal densities for a sampling algorithm, 
a problem which has posed significant problems
for MCMC in the case of multi-modal likelihood functions.
} %marking new text

@@

At this point, each $\vec{a}$ in $\{\hat{A}\}$ should have a corresponding symmetric hyperbox
surrounding it described by the corresponding 
$\vec{a}_\text{max}(\vec{a})$ and $\vec{a}_\text{min}(\vec{a})$.  
To integrate the posterior probability density, we will assume that all of the
parameter space points in each hyperbox correspond have $\chi^2=\chi^2(\vec{a})$,
{\bf %marking new text
i.e. the $\chi^2$ discovered by APS for the point at the center of the hyperbox.
} %marking new text
The posterior probability associated with each box is then
\begin{eqnarray}
P(\vec{a})&=&\frac{V(\vec{a})}{P_\text{tot}}\times\exp\left[-\chi^2(\vec{a})/2\right]
\label{eqn:posterior}\\
P_\text{tot}&\equiv&\sum_{\vec{a}} V(\vec{a})\times\exp\left[-\chi^2(\vec{a})/2\right]
\end{eqnarray}
where $V(\vec{a})$ is the hypervolume of the hyperbox surrounding $\vec{a}$.
To plot the $(1-\alpha)\%$ credible limit, one need only plot the points in hyperspace that
enclose the lowest-$\chi^2$ hyperboxes containing $(1-\alpha)\%$ of the total probability.
We test this algorithm on a function with an exactly known credible limit
in Section \ref{sec:toy} (see Figures \ref{fig:toyBayes2}-\ref{fig:toyBayes4}) 
and on a real, physical dataset in Section \ref{sec:wmap} (see Figure \ref{fig:contours_bayes}).  
In  Section \ref{sec:toy}, we find that our results agree with the known credible limit.
{\bf %marking new text
In Section \ref{sec:wmap} we find that APS correctly identifies the region and general
shape of the Bayesian credible limits, however, with significant noise.  The Bayesian
credible limits found by APS are systematically larger than the Bayesian credible
limits found by MCMC.  This motivates our claim in the introduction that APS may be
principally useful for identifying regions of high posterior density on largely unknown
likelihood functions, allowing users to more shrewdly set priors for sampling
algorithms like MCMC.
}%marking nex text

@@

In Figure \ref{fig:contours_bayes}, we plot the 95\% Bayesian credible limit discovered
by APS after 50,000 evalutations of $\chi^2$ and determined by the method outlined in
Section \ref{sec:bayes}.  Unfortunately, we do not see here the convergence to the true
credible limits that we say in Figures \ref{fig:toyBayes2}-\ref{fig:toyBayes4}.  While
it does appear that APS has correctly identified the locations and general shapes of the
true credible limits, it has, in most cases, drawn credible limits that are larger than
those found by MCMC.  This should not be surprising.  APS was designed with Frequentist
confidence limits in mind, and Figures \ref{fig:contours_gauss}-\ref{fig:contours_br}
seem much better converged than Figure \ref{fig:contours_bayes}.  However, it does seem
that APS might be useful in cases where little prior knowledge is known about $\vec{\theta}$
and users need to be confident that they have discovered all of the low-$\chi^2$
regions in a particular parameter space.  Once APS has been used to identify rough
Bayesian credible limits, more detailed algorithms (like MCMC) can be run using the prior
knowledge gained by APS to refine the credible limits.

@@

In Section \ref{sec:toy} we indicated that, consistent with the findings
of Bryan {\it et al}. (2007), APS is more robust
than MCMC against multi-modal likelihood functions (at least on the idealized
cartoon $\vec{\theta}\rightarrow\chi^2$ functions considered).
{\bf %marking new text
In Section \ref{sec:wmap}, we showed that the presence of realistic noise
does not interfere with APS's ability to derive Frequentist confidence limits.
}%marking new text
However, the Bayesian credible
limits yielded by APS remain noisy and imprecise.  We believe that this performance
can still be of use to the community, especially when trying to evaluate proof-of-concept
parameter constraints (perhaps for forecasting future parameter constraints)
with very expensive $\vec{\theta}\rightarrow\chi^2$ functions on poorly-understood
parameter spaces.  Certainly, for users interested in plotting Frequentist confidence 
limits, APS is a viable option.
